### 2016-07-05 text/abstraction_relation.tex
Sketching of abstraction relation formalities.

We also define a utility function getRules: PC $\rightarrow$ Set(Rules). This function accepts a path condition and returns the rules which were involved in its construction. This information will be used to report the status of contract proof. Note that operationally, this is done by storing the names of the rules within the path condition.

@@

Finally, we define a function getComponents: PC $\rightarrow$ Set(Typed Graph). This function is used for the abstraction relation, as it returns a set of subsets from the match graph of the path condition. These components are collections of the match graph of rules that have been overlaid on top of each other. These components must be matched as a unit, as their elements must be injectively matched. However, elements from different components may match onto the same target vertex in the abstraction relation. More information on how the components are constructed will be provided in the section on path condition construction. \bentley{This is not elegant at all...}

@@

Let us start by formally defining the notion of abstraction of a transformation
execution by a path condition.

LOOK AT COMPONENTS IN PC MATCH GRAPH AS ELEMENTS AND TRACE LINKS

THEN LOOK AT OUTPUT ELEMENTS AND TRACE LINKS IN IOM

REDO STRUCTURES TO ONLY BE ONE VERTEX AND EDGE SET, AND HAVE SUBSETS FOR MATCH, APPLY

@@

\noindent\makebox[\linewidth]{\rule{\linewidth}{0.4pt}}
The PC's NAC does not match onto the iom's Input

\begin{multline}
\label{eq:abstr_no_nac}
\nexists \big(\mathit{PC}_{NAC} \stackrel{inj}{\vartriangleleft} \mathit{iom}_{Input}^*\big)
\end{multline}

\noindent\makebox[\linewidth]{\rule{\linewidth}{0.4pt}}

Each component in the PC:
Elements matches on the iom's Input
Traceability links can be found as well

\begin{multline}
\label{eq:abstr_input_output}
\big(\forall \mathit{comp} \in \mathit{getComponents(PC)} \vartriangleleft Input^* : \mathit{comp} \stackrel{inj}{\vartriangleleft} \mathit{iom}_{Input}^*\big)
\end{multline}

\noindent\makebox[\linewidth]{\rule{\linewidth}{0.4pt}}

Elements in the iom's Output graph must surjectively match on the Output graph of the PC
Trace links in iom can be found in PC

\begin{multline}
\label{eq:abstr_apply_surjective_match}
\mathit{iom}_{Output} \stackrel{surj}{\vartriangleleft} \mathit{PC}_{Output}
\end{multline}

\noindent\makebox[\linewidth]{\rule{\linewidth}{0.4pt}}

%\begin{multline}
%\label{eq:abstr_trace}
%
%\forall \mathit{PCTrace} \in \mathit{PC}_{trace} :
%
%\exists \mathit{IOMTrace} | \mathit{IOMTrace}
%
%\big(\forall trc\in Components(pc|_{trace})\;.\;trc\vartriangleleft ex\big)\;\land\;\\
%\big(\forall e'\in E'\, \exists e\in E\;.\;\tau'(e')=trace \Longrightarrow \tau(e)=trace\big)
%\end{multline} 
% where for all $rl,rl'\in Rules$, if $rl\cong rl'$ then there exist two injective typed graph homomorphisms $f$ and $g$ such that $Match(rl)\stackrel{f}{\vartriangleleft} Input^*$, $Match(rl)\stackrel{g}{\vartriangleleft} Input^*$ and $CoDom(f)\neq CoDom(g)$.
% $/pc/ = Apply_{ex}\cup ex|_{trace}$ is the output part of $pc$, including symbolic traceability links.

\reviewer{Definition 28: When you define the abstraction of an execution
  by a path condition, you require two conditions (separated by an
  and). I have the feeling that the morphisms used in both
  conditions should somehow be compatible to each other. However,
  your triangle notation indicates only the existence of a morphisms.
  How do you express that these morphisms somehow "agree"?}

@@

@@

To understand the abstraction relation in \cref{def:abstraction_pc_ex}, recall
that during the construction of a transformation execution rules are matched
injectively in the input model. This information is present in the first
condition of the abstraction relation (Proposition~\ref{eq:abstr_input_output})
via the injective typed graph homomorphism between the match part of the copies
of rules ``glued'' onto the path condition and the containment transitive
closure of the input part of the transformation execution. This relation
enforces the fact that certain parts of the execution were found, or matched, by
certain parts of the path condition.
On the other hand, the surjection from the output of the execution towards the
apply part of the path condition models the fact that the output of the
execution has been completely built by instantiating the apply parts of the
rules contained in the path condition.

>>>>

We also define a utility function getRule: V $\rightarrow$ Rules. This function accepts a vertex from the path condition and returns the rule which created that vertex. This function will be used during the matching procedure of the abstraction relation, as elements from the same rule cannot match over the same target element. Note that this function allows production of the set of rules which are represented by the path condition. This information will be used to report the status of contract proof.

@@

%Finally, we define a function getComponents: PC $\rightarrow$ Set(Typed Graph). This function is used for the abstraction relation, as it returns a set of subsets from the match graph of the path condition. These components are collections of the match graph of rules that have been overlaid on top of each other. These components must be matched as a unit, as their elements must be injectively matched. However, elements from different components may match onto the same target vertex in the abstraction relation. More information on how the components are constructed will be provided in the section on path condition construction. \bentley{This is not elegant at all...}

@@

\subsubsection{Formal Definition}
\label{subsubsec:abstraction_relation_formal}
Let us start by formally defining the notion of abstraction of a transformation execution by a path condition.

@@

@@

\paragraph{Condition 1:}
The PC's NAC does not match onto the iom's Input

\begin{equation}
\label{eq:abstr_no_nac}
\nexists \big(\mathit{PC}_{NAC} \stackrel{inj}{\vartriangleleft} \mathit{iom}_{Input}^*\big)
\end{equation}

\paragraph{Condition 2:}

The PC's Input matches onto the iom's Input. However, elements from the same rule in the PC cannot match over the same iom element. Note that this match function \textit{f} is not injective.

\begin{equation}
\label{eq:abstr_input_output}
\exists \big(\mathit{PC}_{Input} \vartriangleleft \mathit{iom}_{Input}^*
where \forall v_1, v_2 \in V_{PC-Input} :
f(v_1) = f(v_2) \veebar getRule(v_1) = getRule(v_2) 
\big)
\end{equation}

\paragraph{Condition 3:}

Elements in the iom's Output graph must surjectively match on the Output graph of the PC

\begin{equation}
\label{eq:abstr_apply_surjective_match}
\mathit{iom}_{Output} \stackrel{surj}{\vartriangleleft} \mathit{PC}_{Output}
\end{equation}

\paragraph{Condition 4:}

Traceability links in the PC must injectively match onto links in the transformation execution. This is to ensure that element creation represented in the path condition reflects actual rule execution present in the transformation execution.

TODO: Define the InjMatch we need here

\begin{equation}
\label{eq:abstr_links_injective}
\mathit{PC}_{trace} InjMatch \mathit{iom}_{trace}
\end{equation}

\paragraph{Condition 5:}

Traceability links in the transformation execution must surjectively match onto links in the transformation execution. This case is to ensure that there is no element creation in the transformation execution that is not captured by the path condition.

TODO: Define the SurjMatch we need here

\begin{equation}
\label{eq:abstr_links_surjective}
\mathit{PC}_{trace} SurjMatch \mathit{iom}_{trace}
\end{equation}

\paragraph{Condition 6:}

The morphisms used in the above conditions must match.

TODO: Define how the morphisms agree.

\begin{equation}
\label{eq:abstr_morphisms_agree}
\mathit{PC}_{trace}
\end{equation}

@@

\subsubsection{Explanation}

To understand the abstraction relation in \cref{subsubsec:abstraction_relation_formal}, recall
that during the construction of a transformation execution rules are matched
injectively in the input model. This information is present in the first
condition of the abstraction relation (Proposition~\ref{eq:abstr_input_output})
via the injective typed graph homomorphism between the match part of the copies
of rules ``glued'' onto the path condition and the containment transitive
closure of the input part of the transformation execution. This relation
enforces the fact that certain parts of the execution were found, or matched, by
certain parts of the path condition.
On the other hand, the surjection from the output of the execution towards the
apply part of the path condition models the fact that the output of the
execution has been completely built by instantiating the apply parts of the
rules contained in the path condition.

### 2016-07-05 text/formal_background.tex
Sketching of abstraction relation formalities.

\begin{definition}{Typed Graph\\}
\label{def:typed_graph}
A typed graph is a 6-tuple $\langle V,E,(s,t),\tau, \VT, \ET\rangle$ where:
\begin{itemize}
\item $V$ is a finite set of vertices
\item $E$ is a finite set of directed edges connecting the vertices $V$
\item $(s,t)$ is a pair of functions $s: E\rightarrow V$ and $t: E\rightarrow V$ that respectively provide the source and target vertices for each edge in the graph
\item Function $\tau:V\cup E\rightarrow \VT \cup \ET$ is a typing function for the elements of $V$ and $E$, where $\VT$ and $\ET$ are disjoint finite sets of vertex and edge type identifiers and $\tau(v)\in \VT$ if $v\in V$ and $\tau(e)\in \ET$ if $e\in E$
\item Edges $e\in E$ are noted $v\xrightarrow{e} v'$ if $s(e)=v$ and $t(e)=v'$, or simply e if the context is unambiguous
\item The set of all typed graphs is called $\textsc{TG}$
\item We define the empty graph to be a graph with $V = \emptyset$, $E = \emptyset$, and the other elements to be empty functions
\end{itemize}
\end{definition}

@@

@@

%, where $\tau(v_1)=\tau'(f(v_1))$, $\tau(v_2)=\tau'(f(v_2))$ and also $\tau(e)=\tau(e')$.
\item The domain of $f$ is noted $Dom(f)$ and the co-domain of $f$ is noted $CoDom(f)$.
\end{itemize}  
\end{definition}

 When an \emph{injective} typed graph homomorphism $f$ exists from $g$ onto $g'$, we write $g \stackrel{inj}{\blacktriangleleft} g'$. When a \emph{surjective} typed graph homomorphism $f$ exists from $g$ onto $g'$ we write $g \stackrel{surj}{\blacktriangleleft} g'$.

@@

Our technique also requires a form of graph homomorphism that primarily focuses on matching the edges of the graphs. This homomorphism will be an injective\footnote{For a terminology refresh, \textit{injective matching} means there is a one-to-one correspondence between the source and target vertices. \textit{Surjective matching} means that all target vertices must be matched by at least one source vertex.} match regarding the edges, but note that a particular vertex in the source graph may match onto multiple vertices in the target graph. \bentley{What's the term for this?}

>>>>

\begin{definition}{Typed Graph\\}
\label{def:typed_graph}
A typed graph is a 6-tuple $\langle V,E,(s,t),\tau, \VT, \ET\rangle$ where:
\begin{itemize}
\item $V$ is a finite set of vertices
\item $E$ is a finite set of directed edges connecting the vertices $V$
\item $(s,t)$ is a pair of functions $s: E\rightarrow V$ and $t: E\rightarrow V$ that respectively provide the source and target vertices for each edge in the graph
\item Function $\tau:V\cup E\rightarrow \VT \cup \ET$ is a typing function for the elements of $V$ and $E$, where $\VT$ and $\ET$ are disjoint finite sets of vertex and edge type identifiers and $\tau(v)\in \VT$ if $v\in V$ and $\tau(e)\in \ET$ if $e\in E$
\item Edges $e\in E$ are noted $v\xrightarrow{e} v'$ if $s(e)=v$ and $t(e)=v'$, or simply e if the context is unambiguous
\item The set of all typed graphs is called $\textsc{TG}$
\item We define the empty graph to be a graph with all elements to be empty functions or sets
\end{itemize}
\end{definition}

@@

We also define a function getEdges: \textit{ET} $\rightarrow$ Set(Edges) to return all edges of a certain edge type.

@@

%, where $\tau(v_1)=\tau'(f(v_1))$, $\tau(v_2)=\tau'(f(v_2))$ and also $\tau(e)=\tau(e')$.
%\item The domain of $f$ is noted $Dom(f)$ and the co-domain of $f$ is noted $CoDom(f)$.
\end{itemize}  
\end{definition}

 When an \emph{injective}\footnote{For a terminology refresh, \textit{injective matching} means there is a one-to-one correspondence between the source and target vertices. \textit{Surjective matching} means that all target vertices must be matched by at least one source vertex.} typed graph homomorphism $f$ exists from $g$ onto $g'$, we write $g \stackrel{inj}{\blacktriangleleft} g'$. When a \emph{surjective} typed graph homomorphism $f$ exists from $g$ onto $g'$ we write $g \stackrel{surj}{\blacktriangleleft} g'$.

@@

Our technique also requires a form of graph homomorphism that primarily focuses on matching the edges of the graphs. This homomorphism will be an injective match regarding the edges, but note that a particular vertex in the source graph may match onto multiple vertices in the target graph. \bentley{What's the term for this?}
